Formal Verification Framework Based on Symbolic Execution for Smart Contract

### 2023-12-06 papers/static-checker/ecoop2024/ecoop2024.tex
update rules

\title{Formal Verification Tool Based on Symbolic Execution for Smart Contract} %TODO Please add

@@

\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model}
\subsection{System Model}
Michelson is the low-level smart contract language employed in the Tezos blockchain platform. It is characterized as a stack-based language, featuring high-level data types and primitives and integrates strict static type checking into its stack-based execution mode.

The processing and manipulation of data during the execution of smart contracts are facilitated using a stack. A stack  can be written:
\begin{itemize}
\item[]  \EMPTYSTACK\ for the empty stack.
\item[]  \HEAD\ \STACKCONCAT\ \NTAIL\ for the stack whose first value has type (\HEAD) and queue has stack type (\NTAIL).
\end{itemize}

Each piece of data in Michelson is assigned a specific type, and the language imposes restrictions preventing programmers from introducing custom types. Consider a stack denoted as $\STACK = (\TermOne, \TYF) \STACKCONCAT (\TermTwo, \TYS) \STACKCONCAT \DOT
\STACKCONCAT \EMPTYSTACK$ where elements are types paired with terms.

A Michelson program is a sequence of instructions executed in a sequential manner. Each instruction takes, as input, the stack produced by the preceding instruction and modifies it for the subsequent one. This sequential flow ensures that each instruction builds upon the state left by its predecessor in the program. Let $\INSTRUCTION = \InstructionOne; \InstructionTwo; \DOT; \InstructionN$ represent a sequence of instructions.

While undergoing symbolic execution,  a symbolic execution state assumes different values based on the constraints of symbolic values. Additionally, branching may occur depending on conditions, and both constraints and branching conditions are recorded as predicates. Let \PREDICATE\ symbolize a predicate expressed in conjunction form, capturing the conditions and constraints encountered during the symbolic execution process.
\begin{definition}
A symbolic execution's state is represented by the tuple $\STATE = [\INSTRUCTION, \STACK, \PREDICATE]$. Here, $\SYSTEM = \{\STATEONE, \STATETWO, \DOT, \STATEN \}$ range over sets of  states.
\end{definition}
The Michelson program takes as input a stack that consists of a single pair whose first element is an input value referred to as the parameter, and second element represents the content of the storage space. The program produces a stack as output, containing a single pair. The first element of this pair is the list of internal operations that the program intends to emit, and the second element is the new contents of the storage space.

Define the initial stack as $S_{init}$   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\EMPTYSTACK, where \VPAR\ and \VSTORAGE\ denote terms representing the parameter and the storage, and \TYF\ and \TYS\ are their corresponding types. Similarly, designate the final stack as $S_{final}$   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\EMPTYSTACK, where \VOPERATIONLIST\ repesents a operation list.

%\paragraph{Sequence actions.}
%\begin{figure}[h]
%\center
%\includegraphics[width= 10 cm]{life-cycle-1}
%\caption{Life-cycle for the auction contract.}
%\label{fca1}
%\end{figure}

\pagebreak
\subsection{Rules}

Each instruction is defined by a rule on the system model. The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a state, which is non-deterministic  as shown for loop instructions, such as \LOOP\ and \LOOP-LEFT,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

During the execution of symbolic analysis, numberous states are generated. However, certain states may be unreachable from a symbolic initial state. This occurs when the predicate \PREDICATE\ is unsatisfied. Consequently, a system state is disregarded when the predicate \PREDICATE\ is fail.

\begin{mathpar}
\inferrule[]
  { \NEG\ \PREDICATE
  }{
  \{[\INSTRUCTION, \STACK, \PREDICATE]\} \cup \SYSTEM \SystemTrans \SYSTEM}
\end{mathpar}

There are various types of instructions in Michelson, categorized based on their functions. These include control structures, stack manipulation, arithmetic operations, boolean operations, cryptographic operations, and operations on data structures. In our model, we classify them into three main groups: basic instructions, branch instructions, and loop instructions. This categorization reflects how these instructions are handled within our specific model.
\subsubsection{One-step Instructions}
A one-step instruction only effects a system state itself, does not produce a new state, and halts the execution after a single operation. The ABS rule specifies the corresponding instruction that applies to the first element of the stack. This element must have the type of int and returns the positive value of its current value, which is of a natural number type. Another example, the COMPARE instruction compares the two first elements of a stack, both of which have a natural number type. It returns a result of either -1, 0, or 1 depending on the comperations of their values.
%ABS
\begin{mathpar}
\inferrule[ABS]
  {
  }
  {
    [(\ABS; \INSTRUCTION), (\StackOne, \TINT) \STACKCONCAT \STACK,
    \PREDICATE]
    \StateTrans \
    [\INSTRUCTION, (\VariableX, \TNAT) \STACKCONCAT \STACK,
    \PREDICATE \wedge (\StackOne \ge 0 \Rightarrow \VariableX =
    \StackOne) \wedge (\StackOne <0 \Rightarrow \VariableX = -\StackOne)]
 }
\end{mathpar}

% COMPARE
\begin{mathpar}
\inferrule[COMPARE]
  {
  }
  {
    [(\COMPARE ; \INSTRUCTION), (\StackOne, \TNAT) \STACKCONCAT (\StackTwo, \TNAT)
    \STACKCONCAT \STACK, \PREDICATE ]
    \SystemTrans \\
    [\INSTRUCTION, (\VariableX, \TINT) \STACKCONCAT \STACK, \PREDICATE
    \wedge\ (\StackOne > \StackTwo \Leftrightarrow \VariableX = 1)
    \wedge\ (\StackOne = \StackTwo \Leftrightarrow \VariableX = 0) 
    \wedge\ (\StackOne < \StackTwo \Leftrightarrow \VariableX = -1)]
    }
\end{mathpar}
\subsubsection{Multi-step Instructions}
Multi-step instructions require the execution of a sub-sequence of instructions before returning to the main execution. The following rule models the instruction EXEC, which executes a function specified as a sequence of instructions denoted as \INSTRUCTIONONE\ in the rule. This instruction first applies the code of the function to the first element of the stack. It then places the result back into the main execution.
\begin{mathpar}
  \inferrule[EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge Q']}
\end{mathpar}

@@

\subsubsection{Cryptographic Instructions}
Cryptographic instructions in Michelson utilize built-in functions that manage cryptographic features. These functions are symbolically modeled in our framework, specifying their inputs and outputs. For instance, the instruction \SHA\ computes the cryptographic hash of the top of the stack using the SHA-256 cryptographic hash function. The function SHA-256 is symbolically represented as the function  \FSHA\  that takes an input of type \TBYTE\ and returns another byte value, constrained to a length of 256.
%HASH-KEY
\begin{mathpar}
\inferrule[\SHA]
  {
  }
  {[(\SHA; \INSTRUCTION), (\StackOne, \TBYTE) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VariableX, \TBYTE) \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\VariableX\ = \FSHA(\StackOne))]}
\end{mathpar}
\subsubsection{Branch Instructions}
The Michelson language incorporates conditional branching, enabling the execution of specific code segments based on certain criteria. One of the instructions facilitating this is the IF {} {} construct. For example, the \IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO\ instruction allows the creation of branches in execution. It takes two sequences as arguments. Expecting a boolean value at the top of the stack, it consumes this top element. If the boolean is True, it executes the first provided sequence; otherwise, it executes the second sequence.
%IF
\begin{mathpar}
  \inferrule[IF]
  {  
  }{
    \{[(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]\} \cup \SYSTEM 
      \SystemTrans 
    \{[\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]\} \cup \{ [\INSTRUCTIONTWO, \STACK, \PREDICATE\ \Wedge\ \NEG\
   \StackOne]\} \cup  \SYSTEM
  }
\end{mathpar}

The \IFLEFT\ \INSTRUCTIONONE\  \INSTRUCTIONTWO\ instruction expects a or value at the top of the stack, it consumes this top element. If the value is Left, it executes the first provided sequence; otherwise, it executes the second sequence.
%IF-LEFT
\begin{mathpar}
  \inferrule[IF-LEFT]
  {  
  }{
    \{[(\IFLEFT\ \INSTRUCTIONONE\ \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]\}  \cup  \SYSTEM
    \SystemTrans \\
    \{[\INSTRUCTIONONE, (\VariableX, \TYF) \STACKCONCAT\STACK,
    \PREDICATE \wedge (\StackOne\ \EQUAL\ \LEFT\ \VariableX)]\}  \cup  \{  [\INSTRUCTIONTWO, (\VariableX, \TYS) \STACKCONCAT\STACK, \PREDICATE \wedge (\StackOne\ \EQUAL\ \RIGHT\ \VariableX))]\} \cup \SYSTEM
  }
\end{mathpar}

\subsubsection{Loop Instructions}
In Michelson, there are several loop instructions that execute a loop over the data structure, such as ITER, MAP, MEMBER, SIZE and UPDATE on list, set, map or on the stack, such as LOOP and LOOP-LEFT.

%%ITER
\begin{mathpar}
  \inferrule[ITER-empty]
  {  
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, \STACK,  \PREDICATE\ \Wedge\  (\StackOne\ \EQUAL\ \EMPTYLIST)]
  }
\end{mathpar}

Let us consider the \ITER\ instruction (for list), which applies to a stack \STACK, where the top element has type  (\TY\ \TYLIST) and the rest of the stack has type \TYA, returns the result stack of type \TYA. The \ITER\ instruction loops on the datastructure list. 
Typing
\begin{mathpar}
  \inferrule{\JTypeExpr\TEnv{\INSTRUCTION}{\TY : \TYA\ \SRightarrow\ \TYA}
  }{
      \JTypeExpr\TEnv{\ITER\ \INSTRUCTION}{\TYLIST\ \TY : \TYA\ \SRightarrow\ \TYA}
    }
\end{mathpar}

Semantic

\begin{mathpar} 
  \inferrule{\INSTRUCTION\  /  \VariableX\ \STACKCONCAT\ \STACK \SRightarrow\ \STACK'
  }{
    \ITER\ \INSTRUCTION\ / \{ \VariableX\ ; \TAIL\ \} \STACKCONCAT\ \STACK\ \SRightarrow\ \ITER\ \INSTRUCTION\ / \{ \TAIL\ \} \STACKCONCAT\ \STACK' 
    }
\end{mathpar}

If a list \LIST\ is concrete, the intructuion simply loops until the list ends with an empty list construction. Here we consider the synbolic execution of \ITER\ when the list is a sybolic term or variable.

The sequence of intructuions I is a lamda function  (\TLAMBDA\ (\TY\ : \TYA)\ \TYA) applied to the stack of type (\TY\ : \TYA), returning the stack of type \TYA. We can represent I as the function f that takes an element of type \TY\ and a stack of type \TYA\ and returns a stack of type \TYA.

\begin{mathpar}
  f : \TY\  \TYA\ \SRightarrow\ \TYA
\end{mathpar}

\begin{mathpar}
  f  \VariableX\ \STACK =  (\TLAMBDA\ (\TY\ : \TYA)\ \TYA) (\VariableX\  \STACKCONCAT\ \STACK)
\end{mathpar}

or

\begin{mathpar}
  f  \VariableX\ \STACK = I / (\VariableX\  \STACKCONCAT\ \STACK)
\end{mathpar}

Let  
\LIST\ \STACKCONCAT\ \STACK

 be the stack just before applying the instruction ITER I, where \VariableX\ be the head of the list and \STACK\ be the rest of the stack . For first  loop, let  \STACK\ be f (\VariableX, \STACK). \STACK\ is the stack resulting from applying the function f to the element \VariableX\ and the stack \STACK. In other words, the stack \STACK\ is formed by applying the sequence of the instructions I to the stack \VariableX\ \STACKCONCAT\ \STACK.

Let the result of applying the instruction \ITER\ I on the stack \LIST\ ::  \STACK\ be the stack S', then Z is resulted from the \FOLD\ function that applies the function f  to each element of the list \LIST\ and initial value is the stack \STACK.

\begin{mathpar}
S' =  \FOLD\ \STACK\ f   \LIST
\end{mathpar}

>>>>

\title{Formal Verification Framework Based on Symbolic Execution for Smart Contract} %TODO Please add

@@

\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model}
\subsection{System Model}
Michelson serves as the low-level smart contract language for the Tezos blockchain platform. The processing and manipulation of data during the execution of smart contracts are managed using a stack. The representation of a stack can be described as follows:
\begin{itemize}
\item[]  \EMPTYSTACK\ represents the empty stack.
\item[]  \HEAD\ \STACKCONCAT\ \NTAIL\ denotes a stack with the top element as \HEAD\ and the remaining stack as \NTAIL.
\end{itemize}
\begin{figure} []
\begin{align*}
T, U &::= \\
   &\Mid\ \langle \text{comparable type}\rangle \\
   &\Mid\ \text{option}\ \langle\text{type}\rangle \\
   &\Mid\ \text{list}\ \langle\text{type}\rangle \\
   &\Mid\ \text{set}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \langle\text{type}\rangle \\
   &\Mid\ \text{ticket}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{pair}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{or}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{lambda}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{big-map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{bls12-381-g1} \\
   &\Mid\ \text{bls12-381-g2} \\
   &\Mid\ \text{bls12-381-fr} \\
   &\Mid\ \text{sapling-transaction}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{sapling-state}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{chest} \\
   &\Mid\ \text{chest-key} \\
\langle\text{comparable type}\rangle ::= \\
   &\Mid\ \text{unit} \\
   &\Mid\ \text{never} \\
   &\Mid\ \text{bool} \\
   &\Mid\ \text{int}\\
   &\Mid\ \text{nat}\\
   &\Mid\ \text{string}\\
   &\Mid\ \text{chain-id}\\
   &\Mid\ \text{bytes}\\
   &\Mid\ \text{mutez}\\
   &\Mid\ \text{key-hash}\\
   &\Mid\ \text{key}\\
   &\Mid\ \text{signature}\\
   &\Mid\ \text{timestamp}\\
   &\Mid\ \text{address}\\
   &\Mid\ \text{tx-rollup-l2-address}\\
   &\Mid\ \text{option}\ \langle\text{comparable type}\rangle\\
   &\Mid\ \text{or}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle\\
   &\Mid\ \text{pair}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle \DOT \\
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}
Each piece of data in Michelson is assigned a specific type, and the language restricts programmers from introducing custom types. The complete list of types can be found in Figure \ref{fig:type}. To model the system, we use a stack in which each element is a pair consisting of term and its corresponding type. Terms are defined in Figure \ref{fig:term}. The stack is denoted as \STACK\ = (\TermOne, \TYF) \STACKCONCAT\ (\TermTwo, \TYS) \STACKCONCAT\ \DOT\ \STACKCONCAT\ \EMPTYSTACK\ where elements are types paired with terms.

A Michelson program is a sequence of instructions executed in a sequential manner. Each instruction takes, as input, the stack produced by the preceding instruction and modifies it for the subsequent one. Let \INSTRUCTION\ = \InstructionOne; \InstructionTwo; \DOT\  \InstructionN\ represent a sequence of instructions.
During symbolic execution,  the system state evolves, assuming different values based on the constraints of symbolic values. Moreover, branching may occur depending on conditions, and both constraints and branching conditions are recorded as predicates. Let \PREDICATE\ symbolize a predicate expressed in conjunction form, capturing the branching conditions and constraints encountered during the symbolic execution process.
\begin{definition}
A symbolic execution state is represented by the tuple \STATE\ = [\INSTRUCTION, \STACK, \PREDICATE]. Here, \SYSTEM\ = \{\STATEONE, \STATETWO, \DOT, \STATEN \} range over sets of  states.
\end{definition}

The smart contract program takes a stack as input, consisting of a single pair where the first element is an input value referred to as parameter denoted as \VPAR and the second element represents the content of the storage space denoted as \VSTORAGE. The program produces a stack as output, which is a single pair. The first element of this pair is the list of internal operations that the program intends to emit (\VOPERATIONLIST), and the second element is the new contents of the storage space.  Let's define the initial stack as  \SINIT\   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VPAR\ and \VSTORAGE\ denote terms representing the parameter and the storage, and \TYF\ and \TYS\ are their corresponding types. Similarly, designate the final stack as \SFINAL\   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VOPERATIONLIST\ repesents a operation list.
\begin{figure}[]
\begin{align*}
\text{t} &::= \\
   &\Mid\ \langle \text{variable} \rangle \\
   &\Mid\ \langle \text{account constant} \rangle \\
   &\Mid\ \langle \text{int constant} \rangle \\
   &\Mid\ \langle \text{string constant} \rangle \\
   &\Mid\ \langle \text{byte sequence constant} \rangle \\
   &\Mid\ \UNIT \\
   &\Mid\ \NEVER \\
   &\Mid\ \TRUE \\
   &\Mid\ \FALSE \\
   &\Mid\ \PAIR\ \text{t1 t2}\\
   &\Mid\ \LEFT\ \text{t}\\
   &\Mid\ \RIGHT\ \text{t}\\ 
   &\Mid\ \SOME\ \text{t}\\
   &\Mid\ \NONE \\
   &\Mid\ \text{\{t ; ... \}}\\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}\\
   &\Mid\ \{ \langle\text{instruction}\rangle; ... \}   \\
\langle \text{variable} \rangle &::= \\ 
   &\Mid\ \VAR\\
\langle \text{account constant} \rangle &::= \\ 
   &\Mid\ \text{balance} \\
   &\Mid\ \text{amount} \\
   &\Mid\ \text{sender} \\
   &\Mid\ \text{source} \\
   &\Mid\ \text{now} \\
   &\Mid\ \text{level} \\
   &\Mid\ \text{chain-id} \\
   &\Mid\ \text{self}  \\
   &\Mid\ \text{self-address}  \\
   &\Mid\ \text{total-voting-power}  \\
   &\Mid\ \text{voting-power}  \\
\langle\text{instruction}\rangle &::= \\
  &\Mid\ \DROP \\
  &\Mid\ \DROP \langle\text{natural number constant}\rangle \\
  &\text{...}
\end{align*}
\caption{Terms}
\label{fig:term}
\end{figure}
\subsection{Rules}
Each instruction is defined by a rule on the system model. The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a state, which is non-deterministic  as shown for loop instructions, such as \LOOP\ and \LOOP-LEFT,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

During the symbolic execution, numerous states are generated, but some may be unreachable from a symbolic initial state. This happens when the predicate \PREDICATE\ is unsatisfied, leading to the exclusion of a system state where the predicate fails.

\begin{mathpar}
\inferrule[]
  { \NEG\ \PREDICATE
  }{
  \{[\INSTRUCTION, \STACK, \PREDICATE]\} \cup \SYSTEM \SystemTrans \SYSTEM}
\end{mathpar}
Michelson instructions are categorized based on their functions. These include such as control structures, stack manipulation, arithmetic operations, boolean operations, cryptographic operations, and operations on data structures. In our model, we classified them as one-step, multi-step, blockchain, cryptographic, branch, and loop instructions. This categorization reflects how these instructions are handled within our specific model.
\subsubsection{One-step Instructions}
One-step instructions have a localized effect, modifying the system state directly without branching a new state, and they end the execution after a single operation. The ABS rule exemplifies such an instruction, operating on the first element of the stack. This element must be of type int, and the instruction returns the positive value of its current value, treated as a natural number.
%ABS
\begin{mathpar}
\inferrule[ABS]
  {
  }
  {
    [(\ABS; \INSTRUCTION), (\StackOne, \TINT) \STACKCONCAT \STACK,
    \PREDICATE]
    \StateTrans 
    [\INSTRUCTION, (\VariableX, \TNAT) \STACKCONCAT \STACK,
    \PREDICATE \wedge (\StackOne \ge 0 \Rightarrow \VariableX =
    \StackOne) \wedge (\StackOne <0 \Rightarrow \VariableX = -\StackOne)]
 }
\end{mathpar}
Another example is the COMPARE instruction for natural numbers, which compares the first two elements of the stack, both having a natural number type. It returns a result of either -1, 0, or 1 based on the comparison of their values.
% COMPARE
\begin{mathpar}
\inferrule[COMPARE]
  {
  }
  {
    [(\COMPARE ; \INSTRUCTION), (\StackOne, \TNAT) \STACKCONCAT (\StackTwo, \TNAT)
    \STACKCONCAT \STACK, \PREDICATE ]
    \SystemTrans \\
    [\INSTRUCTION, (\VariableX, \TINT) \STACKCONCAT \STACK, \PREDICATE
    \wedge\ (\StackOne > \StackTwo \Leftrightarrow \VariableX = 1)
    \wedge\ (\StackOne = \StackTwo \Leftrightarrow \VariableX = 0) 
    \wedge\ (\StackOne < \StackTwo \Leftrightarrow \VariableX = -1)]
    }
\end{mathpar}
\subsubsection{Multi-step Instructions}
Multi-step instructions involve the execution of a sub-sequence of instructions before returning to the main execution. The rule below models the \EXEC\ instruction, which executes a function specified as a sequence of instructions denoted by \INSTRUCTIONONE\ in the rule. This instruction applies the code of the function to the first element of the stack and then places the result back into the main execution.
\begin{mathpar}
  \inferrule[\EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge Q']}
\end{mathpar}

@@

\subsubsection{Cryptographic Instructions}
Cryptographic instructions in Michelson utilize built-in functions that manage cryptographic features. These functions are symbolically modeled in our framework, specifying their inputs and outputs. For instance, the instruction \SHA\ computes the cryptographic hash of the top of the stack using the SHA-256 cryptographic hash function. The function SHA-256 is symbolically represented as the function  \FSHA\  that takes an input of type \TBYTE\ and returns another byte value, constrained to a length of 256.
%HASH-KEY
\begin{mathpar}
\inferrule[\SHA]
  {
  }
  {[(\SHA; \INSTRUCTION), (\StackOne, \TBYTE) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VariableX, \TBYTE) \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\VariableX\ = \FSHA(\StackOne))]}
\end{mathpar}
\subsubsection{Branch Instructions}
The Michelson language incorporates conditional branching, enabling the execution of specific code segments based on certain criteria. One of the instructions facilitating this is the \IF\ \{\} \{\} construct. For example, the \IF\ \{ \INSTRUCTIONONE\ \} \{ \INSTRUCTIONTWO\  \} instruction allows the creation of branches in execution. It takes two sequences as arguments. Expecting a boolean value at the top of the stack, it consumes this top element. If the boolean is True, it executes the first provided sequence; otherwise, it executes the second sequence.
%IF
\begin{mathpar}
  \inferrule[IF-true]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]
  }

Another branch instruction is the IF-LEFT instruction, which is used for conditional execution based on the left or right branch of a tagged union (sum) type. 
%IF-LEFT-LEFT
\begin{mathpar}
  \inferrule[IF-LEFT-left]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\ \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONONE, (\VariableX, \TYF) \STACKCONCAT\STACK,
    \PREDICATE \wedge (\StackOne\ \EQUAL\ \LEFT\ \VariableX)]
  }
\end{mathpar}

%IF-LEFT-RIGHT
\begin{mathpar}
  \inferrule[IF-LEFT-right]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONTWO, (\VariableX, \TYS) \STACKCONCAT\STACK, \PREDICATE \wedge (\StackOne\ \EQUAL\ \RIGHT\ \VariableX))]
  }
\end{mathpar}
These above rules collectively define the behavior of the IF-LEFT instruction based on the left or right branch of a tagged union type. This rule applies when the top element of the stack is a tagged union of type \TOR\ \TYF\ \TYS (either of type \TYF or \TYS), and the IF-LEFT instruction is followed by instructions \INSTRUCTIONONE\ and \INSTRUCTIONTWO. The first rule apply with the condition that the top element of the stack is of the form Left \VariableX\ (indicating the left branch of the union), it transitions to execute the instructions \INSTRUCTIONONE\ with the stack modified accordingly and the predicate is updated to include the condition that the top element of the stack is equal to Left \VariableX.  Similerly, The second rule apply with the condition that  the top element is of the form Right \VariableX (indicating the right branch of the union), it transitions to execute the instructions \INSTRUCTIONTWO with the stack modified accordingly and the predicate is updated to include the condition that the top element of the stack is equal to Right \VariableX.

\subsubsection{Loop Instructions}
Loops represent fundamental and powerful programming constructs. In computer programming, a loop is an instruction that iteratively executes a set of statements until a specific condition is met. In a loop structure, the program poses a question, and if the answer necessitates action, the associated code is executed. This questioning and execution process repeats until the condition is no longer met. Each iteration of this cycle is referred to as an iteration of the loop. In Michelson, various loop instructions facilitate the iteration over data structures, allowing for the manipulation and traversal of different data structures within the language. These loop instructions, including ITER, MAP, MEMBER, UPDATE, LOOP, and LOOP-LEFT, are designed for operations on lists, sets, maps, and the stack itself. Similar to other programming languages, Michelson categorizes loop instructions into two main types:
\begin{itemize}
\item[] "For" loops: there are sequence-based loops, such as ITER.
\item[] "While" loops: these are predicates-based loops, illustrated by LOOP and LOOP-LEFT.
\end{itemize}
To illustrate the structures of  loops in Michelson, we can draw a parallel with OCaml syntax:
\begin{verbatim}
while <loop_condition> do
   <loop_body>
done
\end{verbatim}
This "while" loop structure involves a loop condition that is evaluated at the beginning of each iteration. If the condition holds true, the loop body is executed, and the process repeats until the condition becomes false.  In Michelson, this pattern is evident in instructions like LOOP and LOOP-LEFT, which rely on predicate-based conditions for their iteration.
\begin{verbatim}
LOOP {<loop-body>}
\end{verbatim}
Additionally, Michelson features a "For" loop, where the index runs from the initial value to the final value:
\begin{verbatim}
for index = <initial_value> to <final_value> do
  <loop_body>
done
\end{verbatim}
This is represented by the ITER instruction:
\begin{verbatim}
ITER {<loop-body>}
\end{verbatim}
The fundamental concept underlying the definition of a loop is its loop condition, which is the predicate that must hold true for the loop to continue execution. In the case of while loops, the  loop condition is explicitly embedded in the loop syntax. In the case of for loops, the loop condition is more implicit. It can be rewritten as a while loop to make the  loop condition explicit. This "For" loop structure can also be expressed as a "while" loop:
\begin{verbatim}
index = 0
while index < length(sequence) do
  <loop_body>
  index += 1
done
\end{verbatim}
Here, the implicit loop condition asserts that the index of our sequence element is less than the length of the sequence.

It's worth noting that, unlike imperative-style loops, Michelson does not explicitly contain their  loop condition. Instead, the  loop condition is established:
\begin{itemize}
 \item In the expressions preceding the loop for the first time entering the loop.
 \item Inside the loop body for subsequent iterations.
\end{itemize}

The loop instructions in Michelson are applied to a stack, and the loop condition is checked with the first element of the stack. For a "while" loop the first element of the stack must be a boolean value and for the "for" loop, it must be a sequence data that the loop support, such as list, set and map. If we denote the function to retrieve the first element of a stack as "pop," the "while" loop instruction can be expressed as:
\begin{verbatim}
while pop(S) do
   loop_body
done
\end{verbatim}
and the "for" instruction is expressed as
\begin{verbatim}
index = 0
while index < length(pop(S)) do
  <loop_body>
  index += 1
done
\end{verbatim}

Certainly, let's continue exploring how these loop constructs are modeled in the symbolic execution. 
\paragraph {\ITER}
Consider the \ITER\ instruction for lists, which operates on a stack \STACK. Here, the top element of the stack has type (\TY\ \TYLIST), and the rest of the stack has type \TYA. The instruction returns a result stack of type \TYA. The \ITER\ instruction iterates over the data structure list. The corresponding typing rule is as follows:
\begin{mathpar}
  \inferrule{\JTypeExpr\TEnv{\INSTRUCTION}{\TY : \TYA\ \SRightarrow\ \TYA}
  }{
      \JTypeExpr\TEnv{\ITER\ \INSTRUCTION}{\TYLIST\ \TY : \TYA\ \SRightarrow\ \TYA}
    }
\end{mathpar}
This typing rule ensures that if the loop body (\INSTRUCTION) has the type \TY\ : \TYA\ \SRightarrow\ \TYA, then the entire \ITER\ instruction, when applied to a stack with the top element of type (\TY\ \TYLIST) and the rest of the stack of type \TYA, produces a result stack of type \TYA. This facilitates symbolic iteration over a list while maintaining compatibility with further symbolic execution.

We delve into the symbolic execution of \ITER\ when the list is a symbolic term.
%%ITER
\begin{mathpar}
  \inferrule[ITER-empty]
  {  
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, \STACK,  \PREDICATE\ \Wedge\  (\StackOne\ \EQUAL\ \EMPTYLIST)]
  }
\end{mathpar}
This rule handles the case where the list is empty, and thus the \ITER\ instruction does not iterate. The resulting state transitions to the next instruction with an updated stack that is actually the same as before entering the \ITER\ intruction and an augmented predicate \PREDICATE $\Wedge$ (\StackOne \EQUAL \EMPTYLIST).
\begin{mathpar}
  \inferrule[ITER-nonempty]
  {  [\ITER,  (\HEAD, \TY) \STACKCONCAT\STACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK,  \STACK', Q']
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE \wedge Q] \StateTrans \\ [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), \{\STAIL\}\ \STACKCONCAT\STACK',  \PREDICATE \wedge Q' \Wedge  ( \StackOne\ \EQUAL\ \{\HEAD; \STAIL \}) ]
  }
\end{mathpar}
This rule handles the case where the list is non-empty. It symbolically executes the instructions \INSTRUCTIONONE\ within the \ITER\ block with the head of the list (\HEAD) removed from the stack and a subpart of the predicate Q. The resulting state transitions to a new stack (\STACK') and an updated predicate (\PREDICATE\ $\Wedge$ Q' $\Wedge$ (\StackOne \EQUAL \{\HEAD;\STAIL\}), where \{\STAIL\} represents the rest of the list.
In summery, These rules capture the symbolic execution behavior of the \ITER\ instruction, considering both the cases where the list is empty and non-empty.
